\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsthm,amsmath,latexsym,amssymb,amsmath}
\usepackage{fancyhdr,hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{marvosym}
\usepackage{enumitem}
\setlist{nolistsep,leftmargin=*}

\usepackage{titlesec}
\newcommand*{\justifyheading}%{\raggedleft}

\titleformat{\section}
  {\normalfont\large\bfseries%\justifyheading
  \color{Sepia}
  }{\thesection}{1em}{}

\titleformat{\subsection}
  {\normalfont\large\bfseries}{\thesubsection}{1em}{}
\titlespacing*{\section}
{0pt}{6pt}{4pt}
\titlespacing*{\subsection}
{0pt}{6pt}{0pt}


\setlength{\parindent}{0pt}
\setlength{\parskip}{0pt}
\setlist[enumerate]{itemsep=3pt}

%\usepackage{xcolor}
%\usepackage{sectsty}
%\chapterfont{\color{blue}}  % sets colour of chapters
%\sectionfont{\color{indigo}}

\title{
{\textbf{Arka Ghosh}}
%\smallskip
%
%\normalsize
%%\Letter\ arkasparko@gmail.com\\
%\Letter\ a.ghosh@uw.edu.pl\\
%\Letter\ arka.ghosh@u-bordeaux.fr\\
%
%\smallskip
%
%\normalsize
%\Mundus\ https://www.mimuw.edu.pl/\textasciitilde arka/
}
\author{}
\date{}
%\raggedright
\raggedbottom
\begin{document}

\pagenumbering{gobble}

\maketitle
\vspace{-50pt}
\color{black}
\section*{Employment}
\hrule
\smallskip

\textbf{Postdoctoral researcher} in LaBRI(Bordeaux, France)
\hfill January 2026 - current\\
Guided by Nathana\"{e}l Fijalkow.
\medskip

\textbf{Research engineer} in LaBRI(Bordeaux, France) \hfill October 2024 - December 2025\\
Guided by Nathana\"{e}l Fijalkow.

%
\section*{Education}
\hrule
\medskip
\textbf{PhD in Computer Science} from University of Warsaw
\hfill October 2020 - November 2025\\
Guided by S\l{}awomir Lasota, Piotr Hofman.
Awarded grade : passed with distinction.
\medskip

\textbf{Master of Science} from
Indian Institute of Science, Bangalore.\hfill August 2019 - July 2020\\
Major : Mathematics. Awarded grade : first class with distinction.
\medskip

\textbf{Bachelor of Science} from
Indian Institute of Science, Bangalore.\hfill July 2015 - July 2019\\
Major : Mathematics. Awarded grade : first class with distinction.
%
\section*{Participation in Research Grants}
\hrule
\smallskip
\begin{enumerate}
\item[$\diamond$]
From Ocotober 2020 till December 2025 I was a \textbf{research engineer} in the \textbf{PEPR IA project SAIF (Safe AI through formal method)}.
From January 2026 I am a \textbf{postdoctoral researcher} in the same project.
\item[$\diamond$]
From 2023 I am the \textbf{principal investigator} of the NCN preludium project
\textbf{Linear algebra in orbit-finite dimension}.
\item[$\diamond$]
From October 2020 till September 2024 I was a doctoral student in the NCN Opus grant \textbf{Data enriched models of computation}.
\end{enumerate}
%
\section*{Awards and Scholarships}
\hrule
\smallskip
%
\begin{enumerate}
    \item[$\diamond$]
    Distinguished paper award in LICS'23 for the paper \emph{Orbit-finite Linear Programming} co-authored with Piotr Hofman and S\l awomir Lasota.

    \item[$\diamond$]
    Awarded IDUB Scholarship for foreign internship in 2022.
    
     \item[$\diamond$]
    Awarded KVPY Scholarship for Bachelors and Masters studies. 
\end{enumerate}

\section*{Research Visits}
\hrule
$ $\\
July-October 2022,
\textbf{University of Antwerp,
Belgium}.
(\emph{funded by scholarship awarded through IDUB program})
%
\section*{Supervisions}
\hrule
\smallskip
%
\begin{enumerate}
\item[$\diamond$]
Summer internship (August - September 2025) of Antoni Wi\'{s}niewski on weighted automata.
Co-supervised with Nathana\"{e}l Fijalkow.
\item[$\diamond$]
Summer internship (July - September 2023) of Micha\l{} Maszkowski on polynomial ideals.
\end{enumerate}
%
\section*{Publications and Preprints}
\hrule
\smallskip


\textbf{Publication practices:} 
In the theoretical computer science community it is customary to write the authors in the alphabetical order of their surnames. We have followed this custom in all of the papers that I have co-authored.
It is also more common to publish in peer-reviewed conferences than in journals.
Conferences are ranked by CORE (https://portal.core.edu.au/conf-ranks/).
\subsection*{Preprints}
%
\smallskip

\begin{enumerate}
\item[$\diamond$]
Arka Ghosh, Aliaume Lopez:
\textbf{Computability of Equivariant Gr\"{o}bner bases}.
Arxiv preprint.
\item[$\diamond$]
Nathana\"{e}l Fijalkow, Arka Ghosh, Roman Kniazev, Guillermo A. P\'{e}rez, Pierre Vandenhove: \textbf{Computing the Reachability Value of Posterior-Deterministic POMDPs}. Arxiv preprint.
\end{enumerate}
%
\subsection*{Publications in Journals}
\smallskip
\begin{enumerate}
\item[$\diamond$]
Micha\"{e}l Cadilhac, Arka Ghosh, Guillermo A. P\'{e}rez, Ritam Raha:
\textbf{Parikh One-Counter Automata}.
Information and Computation. Volume 306.
\item[$\diamond$] Shaull Almagor, Arka Ghosh, Tim Leys, Guillermo A. P\' erez:\\
\textbf{The Geometry of Reachability in Continuous Vector Addition Systems with States.}\\
Information and Computation. Volume 304.
\item[$\diamond$] Arka Ghosh, Piotr Hofman, S\l awomir Lasota:
\textbf{Orbit-finite linear programming.}
Journal of the ACM, Volume 72, Issue 1.
(\textcolor{BrickRed}{\textit{Full version of LICS distinguished paper}})
\end{enumerate}
%
\subsection*{Publications in Peer-Reviewed Conferences}
\smallskip
\begin{enumerate}
\item[$\diamond$]
Arka Ghosh, S\l{}awomir Lasota : \textbf{Equivariant ideals of polynomials}.
LICS 2024.
\item[$\diamond$]
Micha\"{e}l Cadilhac, Arka Ghosh, Guillermo A. P\'{e}rez, Ritam Raha:
\textbf{Parikh One-Counter Automata}.
MFCS 2023.

\item[$\diamond$] Shaull Almagor, Arka Ghosh, Tim Leys, Guillermo A. P\'{e}rez:
\textbf{The Geometry of Reachability in Continuous Vector Addition Systems with States.}MFCS 2023.
\item[$\diamond$]
Arka Ghosh, Piotr Hofman, S\l awomir Lasota:\textbf{Orbit-finite linear programming.}LICS 2023.
(\textcolor{BrickRed}{\textit{Distinguised paper award. Invited for submission to the Journal of the ACM.}})
\item[$\diamond$]
Arka Ghosh, Piotr Hofman, S\l awomir Lasota:
\textbf{Solvability of orbit-finite systems of linear equations.}
LICS 2022.
\end{enumerate}
%
\section*{Presentations in International Conferences with Selection Committee}
\hrule
\smallskip

\textbf{Equivariant Polynomial ideals} (LICS 2024, Highlights 2024).
\textbf{Reachability in Symmetric Petri Nets with Data} (Highlights 2023).
\textbf{Parikh One-Counter Automata} (MFCS 2023).
\textbf{Orbit-finite Linear Programming} (LICS 2023).
\textbf{Orbit-Finite Systems of Linear Equations} (Highlights 2022).
\textbf{Solvability of Orbit-finite Systems of Linear Equations} (Highlights 2021).
%
%\section*{Seminar Presentations}
%\hrule
%\smallskip
%
%\textbf{Equivariant polynomial ideals and their membership problem}.\\
%June 2025, M2F, LaBRI.
%\smallskip
%
%\textbf{Equivariant Polynomial Ideals}.\\
%October 2023, Automata Seminar, MIMUW.
%\smallskip
%
%\textbf{Orbit-finite Systems of Linear Inequalities}.\\
%March 2023, Automata Seminar, MIMUW.
%\smallskip
%
%\textbf{Orbit-finite System of Linear Equations}.\\
%December 2021, Automata Seminar, MIMUW.
%\smallskip

\section*{Teaching Experience}
\hrule
\smallskip

\textbf{Theory of concurrency} (teaching assistant, winter semester 2023/24).
\textbf{Languages, automata and computations} (teaching assistant, summer semester 2022/23),
Languages, automata and computations II (teaching assistant, winter semesters 2021/22 and 2022/23),
\textbf{Computational complexity} (teaching assistant, winter semester 2021/22),
\textbf{Algorithms for Data Science} (teaching assistant, summer semester 2020/21).
\section*{Services}
\hrule
\smallskip

\paragraph*{Committee Membership:}
%
VMCAI 2026 artifact evaluation committee member.
%
\vspace{-10pt}
\paragraph*{External reviewer:}
STACS 2026,
SODA 2026,
CONCUR 2025,
STACS 2025,
ICALP 2024,
MFCS 2023,
CONCUR 2023,
LICS 2022.
%
%
%\section*{Languages}
%\hrule
%\smallskip
%
%\textbf{Human Languages:} Bengali, English, Hindi
%\smallskip
%
%\textbf{Programming Languages:}
%Idris, Sagemath, Haskell, Python, Rust
\end{document}